---
title: 'Analysis of effects of cortical inactivation on vowel discrimination in noise
  using logistic regression  '
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
  html_notebook: default
---

Version History:

* Created: 2021-09-09 Stephen Town

* Ported to notebook: 2022-05-19

Combine data across ferrets and run a generalized linear mixed model to determine the effect of cooling on log-odds of correct response, with subject and test  session as random effects, and cooling as fixed effects

The individual observations are at the trial level, which is important for analysis of error magnitude, but means that the data for trial accuracy (correct/error) is in binary format.


```{r}
rm(list = ls())

suppressPackageStartupMessages( library(dplyr))
suppressPackageStartupMessages( library(tidyverse))

suppressPackageStartupMessages( library(lme4))
suppressPackageStartupMessages( library(DHARMa))
suppressPackageStartupMessages( library(scales))
suppressPackageStartupMessages( library(sjPlot))
```

## 1. Load data

Load data from ferrets with working cooling loops / optic fibers and select specific columns. 

```{r}
data_path = '/home/stephen/Github/Vowel_Discrimination_In_Noise/Results/Vowels_in_Noise/data/analysis'

magnum = read_csv( file.path( data_path, 'F1311.csv'), show_col_types = FALSE)
robin  = read_csv( file.path( data_path, 'F1509.csv'), show_col_types = FALSE)
mimi   = read_csv( file.path( data_path, 'F1706.csv'), show_col_types = FALSE)

cols = c('fNum','treatment',"vowel","Mask","Correct","vowel_level","SNR","CenterReward")
df = rbind(magnum[cols], robin[cols], mimi[cols])

df$fNum <- as.factor(df$fNum)
head(df)
```

```{r}
table(df$fNum)

nrow(df)
```


## 2. Transform

Although not strictly necessary for models without regularization, it is helpful for the interpretation of coefficients if we scale the sound level of vowels between 0 and 1. We then count the number of correct trials in each unique condition, and the total number of trials; again this isn't strictly needed, but count data can be more convenient to work with when inspecting model outputs.

```{r}
df$vowel_level_orig <- df$vowel_level    # Save a copy for reconstruction
df$vowel_level <- rescale(df$vowel_level)

counts = df %>% 
  group_by(fNum, treatment, Mask, vowel, vowel_level, CenterReward) %>%
  summarise(
    nCorrect = sum(Correct),
    nTotal = n(),
    .groups = 'keep'
    ) %>%
  mutate(pCorrect = nCorrect / nTotal * 100)

head(counts)
```



## 3. Fit Model

Fit a logistic regression model with treatment (on/off), mask (clean or with noise) and vowel sound level as factors

```{r}
mdl_1 = glmer(
  cbind(nCorrect, nTotal - nCorrect) ~ 1 + treatment*Mask + vowel_level + CenterReward + (1|fNum), 
  family=binomial, 
  data=counts)

summary(mdl_1)
```

### 3.2. Two two-way interactions

Add interaction between mask and vowel level (predicted)

```{r}
mdl_2 = glmer(
  cbind(nCorrect, nTotal - nCorrect) ~ 1 + treatment*Mask + Mask*vowel_level + CenterReward + (1|fNum), 
  family=binomial, 
  data=counts)

summary(mdl_2)
```

### 3.3. Model 3: Three-way Interaction 

Add full combination of interactions between mask, treatment and level (not sure why this would be interesting *a priori*). Note that the AIC resulting from this model is higher than model 

```{r}
mdl_3 = glmer(
  cbind(nCorrect, nTotal - nCorrect) ~ 1 + treatment*Mask*vowel_level + CenterReward + (1|fNum), 
  family=binomial, 
  data=counts)

summary(mdl_3)
```

## 4. Model Selection

From the three variants we looked at, the AIC criteria would favor model 2 (two two-way interactions), and this seems reasonable, given we would expect both the interaction between mask and sound level, and mask and treatment to be important.


```{r}
anova(mdl_1, mdl_2, mdl_3)
```

```{r}
#tab_model(mdl_2)
```



---


## 5. Check model

### 5.1. Predictions

Use the model to predict the correct response probability. Then convert to correct trial numbers for each unique condition and aggregate over covariates (center reward, vowel level) for plotting.

```{r}
counts$fit <- predict(mdl_2, counts, type="response")

# Count trials for ferret, treatment, mask and sound level
agg_by_level <- counts %>%
  mutate(
    predicted_correct = fit * nTotal
    ) %>%
  group_by(
    fNum, treatment, Mask, vowel_level
    ) %>%          
  summarise(
    nCorrect = sum(nCorrect),
    nTotal = sum(nTotal),
    predicted_correct = sum(predicted_correct),
    .groups = 'keep'
    ) %>%
  mutate(
    pCorrect = nCorrect / nTotal * 100,
    fit_p = predicted_correct / nTotal * 100
  )

# Restore original vowel levels for plotting 
agg_by_level$vowel_level <- rescale(
  agg_by_level$vowel_level,
  to = c(min(df$vowel_level_orig) , max(df$vowel_level_orig)))

# Calculate SNR
agg_by_level$SNR = agg_by_level$vowel_level - 70
```

Plot with vowel level


```{r}
# Change name of clean trials to alter order of the rows in ggplot facets
agg_by_level$Mask[agg_by_level$Mask == 'xClean'] <- 'Clean'
```



```{r}
ggplot( 
  agg_by_level, 
  aes(
    x = vowel_level,
    color = treatment
    )
  ) + 
geom_line(
  aes(y = fit_p)
  ) +
geom_point(
  shape = 19,
  aes(
    y = pCorrect,
    size = nTotal,
    alpha = 0.25
    )
  ) + 
scale_size(range = c(0, 4)) +
labs(
  x = "Vowel Level (dB SPL)", 
  y = "% Correct",
  size='Trials (n)'
  ) + 
scale_color_manual(
  labels = c("Control", "Inactivation"),
  values = c("#3d3d3d", "#37a9a4")
  ) + 
  scale_x_continuous(breaks=seq(50,70,by=10)) +
  scale_y_continuous(breaks=seq(50,100,by=25)) +
theme(
  axis.text = element_text(size = 7.5),
  axis.title = element_text(size = 8),
  strip.text = element_text(size = 8),
  legend.title = element_text(size=8),
  legend.text = element_text(size=7.5)
  #strip.background = element_rect(colour="black", fill="white")
  ) +
facet_grid(
  Mask ~ fNum,
  labeller = labeller(
    Mask = c("Restricted" = "Noise", "Clean" = "Clean"),
    fNum = c("1311" = "F1311", "1509" = "F1509", "1706" = "F1706")
    )
  )


ggsave(filename = "Noise_cooling_predictions.png", width = 12, height = 6.5, units = "cm", dpi=300)
```




### 5.2. Model Checking

```{r}
simulationOutput <- simulateResiduals(fittedModel = mdl_2, n=500, plot = T)
```



