---
title: Effects of spatial separation on vowel discrimination in noise, and it's dependence
  on Auditory Cortex
output:
  pdf_document: default
  html_notebook: default
---

Version History:

* Created: 2021-09-09 Stephen Town

* Ported to notebook: 2022-05-19


```{r}
rm(list = ls())
suppressPackageStartupMessages( library(dplyr))
suppressPackageStartupMessages( library(tidyverse))
suppressPackageStartupMessages( library(lme4))
suppressPackageStartupMessages( library(DHARMa))
suppressPackageStartupMessages( library(scales))
suppressPackageStartupMessages( library(sjPlot))
```


## 1. Load Data

From animals with functional cooling loops (F1311 and F1509) and animals tested with behavior only (F1201, F1203, F1216, F1217). 


```{r}
data_path = '/home/stephen/Github/Vowel_Discrimination_In_Noise/Results/Vowels_Unmasking/data/analysis'

floz = read_csv( file.path( data_path, 'F1201.csv'), show_col_types = FALSE)
virg = read_csv( file.path( data_path, 'F1203.csv'), show_col_types = FALSE)
mini = read_csv( file.path( data_path, 'F1216.csv'), show_col_types = FALSE)
clio = read_csv( file.path( data_path, 'F1217.csv'), show_col_types = FALSE)
magnum = read_csv( file.path( data_path, 'F1311.csv'), show_col_types = FALSE)
robin  = read_csv( file.path( data_path, 'F1509.csv'), show_col_types = FALSE)

cols = c('fNum','treatment',"VowelLocation","SpatialCondition","Atten","F1","Correct")
df = rbind(magnum[cols], robin[cols], floz[cols], virg[cols], mini[cols], clio[cols])

rm(list = c("clio", "floz", "virg", "mini", "robin", "magnum"))
head(df)
```
```{r}
table(df$fNum)

nrow(df)
```


## 2. Transform data

Format subject number as factor and convert binary data to counts.

```{r}
df$fNum <- as.factor(df$fNum)
df$F1 <- as.factor(df$F1)    # First formant = vowel identity

df$vowel_level_orig = 57 - df$Atten
df$vowel_level <- rescale(df$vowel_level_orig)

counts = df %>% 
  group_by(fNum, treatment, VowelLocation, SpatialCondition, F1, vowel_level) %>%
  summarise(
    nCorrect = sum(Correct),
    nTotal = n(),
    .groups = 'keep'
    ) %>%
  mutate( pCorrect = nCorrect / nTotal * 100)

head(counts)
```
## 7. All In One Model

What if we just included everything in the dataset, rather than doing four separate modelling analyses?

```{r}
ACS_counts = subset(counts, SpatialCondition %in% c("colocated", "separated"))

clean_counts = subset(counts, SpatialCondition %in% c("single_speaker"))
clean_counts = rename(clean_counts, clean_pCorrect=pCorrect)
clean_counts = clean_counts %>%
    select(fNum, SpatialCondition, treatment, vowel_level, VowelLocation, F1, clean_pCorrect)

ACS_extended <- merge(
  x = ACS_counts, 
  y = clean_counts, 
  by = c("fNum", "treatment", "VowelLocation", "F1", "vowel_level"), 
  all.x = TRUE, 
  all.y = FALSE)


ACS_extended = ACS_extended %>%
    mutate(clean_pCorrect = clean_pCorrect / 100)

head(ACS_extended, 3)
```
Split data for later testing

```{r}

control_behav = ACS_extended %>%
  filter(treatment == FALSE)


cooled_animals = ACS_extended %>%
  filter(fNum == 1311 | fNum == 1509)

nrow(cooled_animals)

```




### 7.1. Model Fitting

Start with just the main effects

```{r}
ACS_mdl_1 <- glmer(
  cbind(nCorrect, nTotal - nCorrect) ~ SpatialCondition.x + treatment + vowel_level + VowelLocation + (1|fNum) + (0+clean_pCorrect|fNum), 
  data = ACS_extended,
  family = binomial)

summary(ACS_mdl_1)
```
Next, add an interaction between spatial condition and treatment

```{r}
ACS_mdl_2 <- glmer(
  cbind(nCorrect, nTotal - nCorrect) ~ SpatialCondition.x*treatment + vowel_level + VowelLocation + (1|fNum) + (0+clean_pCorrect|fNum), 
  data = ACS_extended,
  family = binomial)

summary(ACS_mdl_2)
```

Add in the interaction between treatment and Attenuation that might capture greater effects of cooling for quieter sounds
```{r}
ACS_mdl_3 <- glmer(
  cbind(nCorrect, nTotal - nCorrect) ~ SpatialCondition.x*treatment + treatment*vowel_level + VowelLocation + (1|fNum) + (0+clean_pCorrect|fNum), 
  data = ACS_extended,
  family = binomial)

summary(ACS_mdl_3)
```

Finally, consider a three-way interaction between cooling, spatial separation and sound level

```{r}
ACS_mdl_4 <- glmer(
  cbind(nCorrect, nTotal - nCorrect) ~ SpatialCondition.x*treatment*vowel_level + VowelLocation + (1|fNum) + (0+clean_pCorrect|fNum), 
  data = ACS_extended,
  family = binomial)

summary(ACS_mdl_4)
```
Repeat model 2, but without the random slope

```{r}
ACS_mdl_5 <- glmer(
  cbind(nCorrect, nTotal - nCorrect) ~ SpatialCondition.x*treatment + vowel_level + VowelLocation + (1|fNum), 
  data = ACS_extended,
  family = binomial)

summary(ACS_mdl_5)
```
What about clean performance as a covariate


```{r}

ACS_mdl_6 <- glmer(
  cbind(nCorrect, nTotal - nCorrect) ~ SpatialCondition.x*treatment + vowel_level + VowelLocation + clean_pCorrect + (1|fNum),
  data = ACS_extended,
  family = binomial)

summary(ACS_mdl_6)
```

```{r}
ACS_mdl_7 <- glmer(
  cbind(nCorrect, nTotal - nCorrect) ~ SpatialCondition.x*treatment + vowel_level + VowelLocation + (1|fNum) + (1|clean_pCorrect), 
  data = ACS_extended,
  family = binomial)

summary(ACS_mdl_7)
```
```{r}
ACS_mdl_8 <- glmer(
  cbind(nCorrect, nTotal - nCorrect) ~ SpatialCondition.x*treatment + vowel_level + VowelLocation + (1|fNum), 
  data = ACS_extended,
  family = binomial)

summary(ACS_mdl_8)
```


```{r}
tab_model(ACS_mdl_8)
```



### 7.2. Model Selection

AIC values


```{r}
anova(ACS_mdl_1, ACS_mdl_2, ACS_mdl_3, ACS_mdl_4)
```
Compare models with same fixed effects, but varying in random effect design

```{r}
anova(ACS_mdl_2, ACS_mdl_6, ACS_mdl_7, ACS_mdl_8)
```




### 7.3. Model Prediction

```{r}
ACS_extended$fit <- predict(ACS_mdl_8, ACS_extended, type="response")


# Summarize performance for ferret, stimulus condition, treatment, and sound level
ACS_agg_by_level <- ACS_extended %>%
  mutate(
    predicted_correct = fit * nTotal
    ) %>%
  group_by(
    fNum, SpatialCondition.x, treatment, vowel_level,
    ) %>%          
  summarise(
    nCorrect = sum(nCorrect),
    nTotal = sum(nTotal),
    predicted_correct = sum(predicted_correct),
    .groups = 'keep'
    ) %>%
  mutate(
    pCorrect = nCorrect / nTotal * 100,
    fit_p = predicted_correct / nTotal * 100,
    group = interaction(fNum, treatment)
  )

# Restore original vowel levels for plotting 
ACS_agg_by_level$vowel_level <- rescale(
  ACS_agg_by_level$vowel_level,
  to = c(min(df$vowel_level_orig) , max(df$vowel_level_orig)))

# Calculate SNR
ACS_agg_by_level$SNR = ACS_agg_by_level$vowel_level - 67
```

Plot

```{r}
ggplot( 
  ACS_agg_by_level, 
  aes(x = SNR, color = SpatialCondition.x)
  ) + 
geom_line(
  aes(y = fit_p)
  ) + 
geom_point(
  shape = 19,
  aes(
    y = pCorrect,
    size = nTotal,
    alpha = 0.25
    )
  ) + 
scale_size(range = c(0, 4))+
scale_color_manual(values=c('#9e0067','#d37000')) + 
labs(
  x = "SNR (dB)", 
  y = "% Correct",
  size='Trials (n)') + 
theme(
  axis.text = element_text(size = 7.5),
  axis.title = element_text(size = 8),
  strip.text = element_text(size = 8),
  legend.title = element_text(size=8),
  legend.text = element_text(size=7.5)
  #strip.background = element_rect(colour="black", fill="white")
  ) +
facet_wrap(
  . ~ group, 
  ncol = 2,
  labeller = labeller(group = 
    c("1201.FALSE" = "F1201",
      "1203.FALSE" = "F1203",
      "1216.FALSE" = "F1216",
      "1217.FALSE" = "F1217",
      "1311.FALSE" = "F1311 (Control)",
      "1509.FALSE" = "F1509 (Control)",
      "1311.TRUE" = "F1311 (Cooled)",
      "1509.TRUE" = "F1509 (Cooled)")
  )
)
 
ggsave(filename = "Unmasking_ACS8_predictions.png", width = 11.2, height = 11.5, units = "cm", dpi=300)

```


```{r}
# Summarize performance for ferret, stimulus condition, treatment
ACS_agg_over_level <- ACS_extended %>%
  mutate(
    predicted_correct = fit * nTotal
    ) %>%
  group_by(
    fNum, SpatialCondition.x, treatment,
    ) %>%          
  summarise(
    nCorrect = sum(nCorrect),
    nTotal = sum(nTotal),
    predicted_correct = sum(predicted_correct),
    .groups = 'keep'
    ) %>%
  mutate(
    pCorrect = nCorrect / nTotal * 100,
    fit_p = predicted_correct / nTotal * 100
  )
```


```{r}

pt = pivot_longer(
      ACS_agg_over_level,
      -c(fNum, SpatialCondition.x, treatment, nTotal, nCorrect, predicted_correct),
      names_to='metric', 
      values_to = 'values')

pt = pt %>% select(fNum, SpatialCondition.x, treatment, metric, values)

pt = pivot_wider(pt, 
    names_from = SpatialCondition.x,
    values_from = values
  )


ggplot(pt, 
    aes(x = colocated, y = separated, color=fNum)
    ) +
geom_point() + 
facet_grid(metric ~ treatment)
```




### 7.4. Model Checking

```{r}
simulationOutput <- simulateResiduals(fittedModel = ACS_mdl_7, n=500, plot = T)
```

Look at the correlation between intercept and slope for random effects


# 8. Running models on subsets of the data to confirm main findings

## 8.1. Control data without cooling - do we still see an effect of spatial unmasking

```{r}
control_model <- glmer(
  cbind(nCorrect, nTotal - nCorrect) ~ SpatialCondition.x + vowel_level + VowelLocation + (1|fNum), 
  data = control_behav,
  family = binomial)

summary(control_model)
```

## 8.2. Effect of cooling with just animals tested with cooling

```{r}
cool_model <- glmer(
  cbind(nCorrect, nTotal - nCorrect) ~ SpatialCondition.x*treatment + vowel_level + VowelLocation + (1|fNum), 
  data = cooled_animals,
  family = binomial)

summary(cool_model)
```